---
title: "Analysis FAA Wildlife strike"
author: "NewMirai | @alangel12407606"
date: "`r Sys.Date()`"
output: pdf_document
---


## Tidy Tuesday

For this 30th edition in 2019 of **Tidy Tuesday** we have a dataset coming from the FAA Wildlife Strike database to look at.
A report on the full dataset can be found [**here**](https://www.faa.gov/airports/airport_safety/wildlife/media/Wildlife-Strike-Report-1990-2017.pdf). 

[**Source**](https://wildlife.faa.gov/)

[**Full data dictionary**](https://wildlife.faa.gov/downloads/fieldlist.xls)


## EDA

First we need to import some basics libraries in order to observe the dataset.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(visdat)
library(skimr)
library(RColorBrewer)
library(glue)
```

Then, we get the data from this [**link**](https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-23/wildlife_impacts.csv)

```{r message=FALSE, warning=FALSE, include=FALSE}
link <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-23/wildlife_impacts.csv"
```


```{r echo=TRUE, message=FALSE, warning=FALSE}
wildlife_impacts <- readr::read_csv(link)
```

We can first, check if we have missing values

```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
visdat::vis_miss(wildlife_impacts,warn_large_data = F)
```

Interesting. It seems that most of the dataset is complete but **17,1%** of missing values is not negligible! We have to be careful about the choices of our variables. 

Let's see a different reprenetation of our data using the `skirmr` library.

```{r echo=TRUE, message=FALSE, warning=FALSE}
skim_to_list(wildlife_impacts)


```


We have a dataset with high dimensions `r dim(wildlife_impacts)`. There is time-series data and geospatial data in this dataset so there is a lot of possibilities for the choice of plot.

## Data wrangling

```{r fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
proportion <- wildlife_impacts %>% 
  select(incident_date,operator,state,damage,time_of_day,atype,time_of_day,phase_of_flt,species,speed) %>% 
  mutate(time_of_day=if_else(is.na(time_of_day),"Unknown",time_of_day),
         damage=case_when(
           damage=="M"~ "Minor",
           damage=="M?"~ "Minor Uncertain",
           damage=="N"~ "None",
           damage=="S"~ "Substantial",
           damage=="D"~ "Destroyed",
           is.na(damage)~ "Unknown Damage"
         ))

nonedmg <- 1.83+2.53+19.63+24.15+37.25

p1 <- proportion %>% 
  ggplot(aes(x=time_of_day,fill=time_of_day))+
  geom_bar(aes(y=(..count..)/sum(..count..)))+
  scale_y_continuous(labels=scales::percent)+
  labs(x="Period of the day",
       y="Relative frequencies",
       title = "Incident per period of day with damage distinction",
       subtitle = glue("{nonedmg} % of incident implies None damage. Good News!!!\n\n"),
       fill="",
       caption = "Author: @alangel12407606 | #TidyTuesday")+
  geom_text(aes( label = paste(round(100*(..count..)/sum(..count..),digits = 2),"%"),
                  y= (..count..)/sum(..count..) ), stat= "count", vjust = -.5,size=3)+
  scale_fill_brewer(type = "qual",palette = "Set3")+
  theme_minimal()+
  facet_wrap(~damage)+
  theme(legend.position = c(0.85,0.25),
        plot.title = element_text(size=18,hjust = .5),
        plot.subtitle = element_text(size=14,hjust = .5),
        strip.background = element_rect(fill = "#f2f2f2"))

p1
```








